---
title: "Homework Assignment 1"
author: "Minglu Wang"
date: 'Assigned: Oct 24, 2020, Due Sun Nov 01, 2020 11:59PM'
output:
  html_document:
    toc: yes
    toc_float: true
  pdf_document:
    toc: yes
---
# Import Data
```{r}
# Import data
nyc <- read.csv("nycflights.csv")

```

## Clean Data

We will now need to clean this data.

```{r}
# Delete the column named "x"
nyc$X <- NULL

# Recode columns as factors as needed
nyc$carrier <- as.factor(nyc$carrier)
nyc$flight <- as.factor(nyc$flight)
nyc$tailnum <- as.factor(nyc$tailnum)
nyc$origin <- as.factor(nyc$origin)
nyc$dest <- as.factor(nyc$dest)

# Re-check summary
str(nyc)
summary(nyc)
```

## Data Exploration

Let's first do some simple exploration of this data. I would like for you to answer each of the questions below with an **inline R code** integrated into the text and not in a separate R code block.

- How many airlines are there? (Hint: `levels` and `length` can be useful here)
```{r}
# Find numbers of airlines

number_of_airline <- length(levels(nyc$carrier))

```
There are `r  number_of_airline ` airlines.

- How many flights there were by the airline with code `OO`? (Hint: `nrow` can be useful here along with logical indexing)
```{r}
# Fine numbers of airline with code "OO"

carrierOO <- nyc[nyc$carrier == "OO",]

nrow(carrierOO)

```
There are `r  nrow(carrierOO) ` flights with the airline code "OO".

- How long is the shortest flight out of any NYC airport? (Hint: `min` can be useful, remember to handle `NA` values)
```{r}
# Find the shortest time

shortest_flight <- min(nyc$air_time, na.rm = TRUE)

```
The shortest flight out of any NYC airport is `r shortest_flight  `min.

- How many flights were there by United Airlines (code: UA) on Jan 12th 2013?
```{r}
# Find the UA flights on Jan 12th 2013

UA01122013 <- nyc[nyc$carrier == "UA" & nyc$year == 2013 & nyc$month == 1 & nyc$day ==12, ]

nrow(UA01122013)

```
There are `r  nrow(UA01122013) ` flights by UA on Jan 12th 2013.

## Arrival Delay

Lets focus on Arrival Delay.

- What was the average arrival delay for all airports and all airlines combined in Jan 2013? 
```{r}
# Find average arrival delay

arr_delay_Jan2013 <- nyc[nyc$year == 2013 & nyc$month == 1, ]

mean(arr_delay_Jan2013$arr_delay, na.rm = TRUE)

```

- Whats was the median arrival delay for all airports and all airlines combined in Jan 2013?
```{r}
# Find median arrival delay

arr_delay_Jan2013 <- nyc[nyc$year == 2013 & nyc$month == 1, ]

median(arr_delay_Jan2013$arr_delay, na.rm = TRUE)

```

Based on your answers to the two questions above, what can you say about the distribution of arrival delays? Provide your answer in a text paragraph form.

  The positive mean arrival delay is larger than the negative median arrival delay, which means there is a fatter tail (more extreme value) on the positive side. The distribution is left-skewed.
  

## Airline Performance

Lets see if all airlines are equally terrible as far as flight arrival delays are concerned. For this question you will have to make sure that airline column is coded as a factor.

- Calculate average arrival delays by airline (Hint: look up the command `tapply`)
```{r}
# Use function tapply and round the number to 2 decimals

round(tapply(nyc$arr_delay, nyc$carrier, mean, na.rm=TRUE), 2)

```

- Draw a Bar Plot of Average Arrival Delays for all the Airlines (Hint: command for making a Bar Plot is simply `barplot`)
```{r}
# Save the matrix above into a new variable and make a bar plot

arr_delay_by_airline <- tapply(nyc$arr_delay, nyc$carrier, mean, na.rm=TRUE)
barplot(arr_delay_by_airline)

```


```{r}
# Sorted barplot

barplot(sort(arr_delay_by_airline, decreasing = TRUE))

```

- Which airline has the highest average arrival delay? Which airline has the smallest average arrival delay? Are there airlines that actually have negative average delay? Provide answer to this question in a text paragraph form using **inline R code**.
```{r}
# Find the highest average arrival delay

max_delay_airline <- names(arr_delay_by_airline[arr_delay_by_airline == max(arr_delay_by_airline)])

# Find the smallest average arrival delay

min_delay_airline <- names(arr_delay_by_airline[arr_delay_by_airline == min(arr_delay_by_airline)])

```

The airline with the highest average arrival delay is : `r  max_delay_airline `.

The airline with the smallest average arrival delay is : `r  min_delay_airline `.

YES. There are airlines that actually have negative average delay as the smallest delay is negative.

## Air Gain

Create a new column named airgain such that airgain = (departure delay - arrival delay) : this is the amount of delay a flight made up while in air.
```{r}
# Create a new column named airgain and name the new table as "nyc2"

airgain <- (nyc$dep_delay - nyc$arr_delay)
nyc2 <- cbind(nyc, airgain)

```

a) Explore airgain data - calculate suitable descriptive statistics and appropriate graphics to better understand this data. This part is open ended - you do what you feel works best for you.
```{r}
# Explore the average airgain by airlines
airgain_by_airline <- round(tapply(nyc2$airgain, nyc2$carrier, mean, na.rm = TRUE), 2)
barplot(sort(airgain_by_airline, decreasing = TRUE))


```

```{r}
max_airgain_airline <- names(airgain_by_airline[airgain_by_airline == max(airgain_by_airline)])

min_airgain_airline <- names(airgain_by_airline[airgain_by_airline == min(airgain_by_airline)])
```
The longest average airgain is `r  max_airgain_airline `.

The shortest average airgain is  `r  min_airgain_airline `.


b) Answer the questions:

  - do airlines actually gain any time when in air on average? 
```{r}
mean(nyc2$airgain, na.rm = TRUE)
```

Airlines actually gain some time when in air on average.


  - Calculate average airgain for different airlines - which airlines do a better job, which do a worse job? 

The same job as I did in part (a) in this section. (same code and plot)

`r  max_airgain_airline ` does the best job, and `r  min_airgain_airline ` does the worst job.


  - Build a graphic  of average airgain for different airlines sorted by airgain. We want to graphically visualize which airlines are good at "catching up" when in air and which airlines actually lose time further after taking off.
  
I think this is the same job as I did in part (a), but I can redo that here.
```{r}
# Average airgain for different airlines sorted by airgain
airgain_by_airline <- round(tapply(nyc2$airgain, nyc2$carrier, mean, na.rm = TRUE), 2)
barplot(sort(airgain_by_airline, decreasing = TRUE))
```

**There is still some problem with my GitHub commit which always remind me the following information.However I have already registered an account. I don't know if there is anywhere to login in to solve this problem.**

Author identity unknown

*** Please tell me who you are.

Run

  git config --global user.email "you@example.com"
  git config --global user.name "Your Name"

to set your account's default identity.
Omit --global to set the identity only in this repository.

fatal: unable to auto-detect email address (got 'dell@DESKTOP-2R89E8R.(none)')
